journal classification

### 2023-12-06 analysis/001_overview.Rmd
Subject figure

publisher_boxplot <- publisher_plot_df |>
  filter(value > 0) |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
  geom_boxplot(outlier.shape = NA, show.legend = FALSE, varwidth = TRUE) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement")) +
  coord_cartesian(ylim = c(0,0.65)) +
  facet_wrap(~publisher, nrow = 1, scales = "free") +
  labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
  theme_minimal() +
    theme(strip.background = element_blank(),
    strip.text.x = element_blank()
    )

@@

>>>>

publisher_boxplot <- publisher_plot_df |>
  filter(value > 0) |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
  geom_boxplot(outlier.shape = NA, show.legend = FALSE, varwidth = TRUE) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement")) +
  coord_cartesian(ylim = c(0,0.65)) +
  facet_wrap(~publisher, nrow = 1) +
  labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
  theme_minimal() +
    theme(strip.background = element_blank(),
    strip.text.x = element_blank()
    )

@@

Next, I focus on the subject distribution.

```{r subject_data_prep}
jn_issnl_ajsc <-
  readr::read_csv(here::here("data", "jn_issnl_ajsc.csv")) |>
  select(issn_l, top_level, subject_area, ajcs)
subject_df <- left_join(jn_ind_jct_all, jn_ind_jct_oa) |>
  mutate(cr_year = as.character(cr_year)) |>
  left_join(ta_active) |>
  mutate(across(everything(), .fns = ~ replace_na(., 0))) |>
  mutate(no_ta = oa - ta_oa) |>
  left_join(jn_issnl_ajsc, by = "issn_l")

top_level_by_year_jn <- subject_df |>
  # with at least one
  filter(oa > 0) |>
  distinct(cr_year, issn_l, jn_all, top_level) |>
  group_by(cr_year, issn_l, top_level) |>
  summarise(n = sum(jn_all)) |>
  mutate(fractions_factor = n / sum(n)) |>
  mutate(fractions = n * fractions_factor) |>
  ungroup() |> 
  group_by(cr_year, top_level) |>
  summarise(articles_n = sum(n),
            articles_frac = sum(fractions)) |>
  arrange(top_level, cr_year)

top_level_by_year_oa_frac <- subject_df |>
   # with at least one
  filter(oa > 0) |>
  select(-ajcs, -subject_area, -esac_publisher) |>
  distinct(cr_year, issn_l, top_level, .keep_all = TRUE) |>
  group_by(cr_year, issn_l, top_level) |>
  summarise(n = sum(jn_all)) |>
  mutate(fractions_factor = n / sum(n)) |>
  select(-n)

top_level_by_year_oa <- top_level_by_year_oa_frac |>
  inner_join(subject_df) |>
     # with at least one
  filter(oa > 0) |>
  select(-ajcs, -subject_area, -esac_publisher) |>
  distinct(cr_year, issn_l, top_level, .keep_all = TRUE) |>
  mutate(across(c(oa, ta_oa, no_ta),  function(x) x * fractions_factor, .names = "frac_{.col}")) |>
  ungroup() |> 
  group_by(cr_year, top_level) |>
  summarise(oa_n = sum(oa),
            oa_frac = sum(frac_oa),
            ta_oa = sum(ta_oa),
            ta_oa_frac= sum(frac_ta_oa),
            no_ta = sum(no_ta),
            no_ta_fractions = sum(frac_no_ta)) |>
  arrange(top_level, cr_year) |>
  inner_join(top_level_by_year_jn)
```

```{r subject_plots}
subject_trend_plot <- top_level_by_year_oa |>
  filter(top_level != "Multidisciplinary") |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  select(cr_year, top_level, contains("frac"),-oa_frac) |>
  pivot_longer(
    cols = c(ta_oa_frac, no_ta_fractions),
    names_to = "type",
    values_to = "value"
  ) |>
  mutate(prop = value / articles_frac) |>
  ggplot(aes(
    gsub("^20", "'", as.character(cr_year)),
    prop,
    fill = type,
    group = type
  )) +
  geom_area() +
  facet_wrap( ~ top_level, nrow = 1) +
  scale_fill_manual(
    "Hybrid OA",
    values = c(
      no_ta_fractions = "#b3b3b3a0",
      ta_oa_frac = "#56B4E9"
    ),
    labels = c(no_ta_fractions = "Other", ta_oa_frac = "Agreement"),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::percent_format(),
    breaks = c(0.05, 0.1, 0.15, 0.2, 0.25)
  ) +
  theme_minimal() +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    axis.text.x = element_blank()
  )

subject_ta_trend_plot <- top_level_by_year_oa |>
  filter(top_level != "Multidisciplinary") |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  mutate(prop = ta_oa_frac / oa_frac) |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
  geom_bar(stat = "identity", fill = "#56B4E9") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25,0.5,0.75), limits = c(0, .75),
                     labels = scales::percent) +
  facet_wrap(~top_level, nrow = 1) +
  theme_minimal() +
  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = NULL) +
  theme(
    axis.text.x = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_blank())
subject_box_trend <- subject_df |>
   # with at least one
  select(-ajcs, -subject_area, -esac_publisher) |>
  distinct(cr_year, issn_l, top_level, .keep_all = TRUE) |>
  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value") |>
  filter(value > 0) |>
  filter(top_level != "Multidisciplinary", !cr_year %in% c("2017", "2023")) |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
  geom_boxplot(outlier.shape = NA, show.legend = FALSE, varwidth = TRUE) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement")) +
  coord_cartesian(ylim = c(0,0.55)) +
  facet_wrap(~top_level, nrow = 1) +
  labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
  theme_minimal() +
    theme(strip.background = element_blank(),
    strip.text.x = element_blank()
    )
```

```{r subject_panel, fig.asp=0.8}
library(patchwork)
design = "E
A
B
C"
wrap_plots(
  E = guide_area(),
  A = subject_trend_plot,
  B = subject_ta_trend_plot,
  C = subject_box_trend,
  design = design,
  guides = "collect",
  heights = c(0, 0.4, 0.3, 0.7)
) &
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.direction = "horizontal",
    panel.grid.minor = element_blank(),
                panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A') 
```

### 2023-12-06 analysis/jn_classification.Rmd
Subject figure

>>>>

```{r side_by_side_subject_publisher}
# normalize publishers
subject_short <- subject_df  |>  mutate(
    publisher = case_when(
      esac_publisher == "Elsevier" ~ "Elsevier",
      esac_publisher == "Springer Nature" ~ "Springer Nature",
      esac_publisher == "Wiley" ~ "Wiley",
      is.character(esac_publisher) ~ "Other"
    )
  ) |>
  mutate(publisher =
           forcats::fct_relevel(
             publisher,
             c("Elsevier", "Springer Nature", "Wiley", "Other") 
           )) |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  # Journal transfers across publishers grouped in publisher
  select(-esac_publisher) |>
# top level
 select(-subject_area, -ajcs) |>
  distinct() |>
  group_by(cr_year, top_level, publisher, issn_l) |>
  summarise(oa = sum(oa), ta_oa = sum(ta_oa), no_ta = sum(no_ta), jn_all = sum(jn_all))

subject_short |>
  filter(top_level != "Multidisciplinary") |>
  ggplot(aes(cr_year, oa / jn_all)) +
  geom_boxplot(varwidth = TRUE) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25,0.5,0.75), limits = c(0, .75),
                     labels = scales::percent) +
  facet_grid(top_level ~publisher)

subject_short |>
  ungroup() |>
  filter(oa > 0) |>
  select(-oa) |>
  filter(top_level != "Multidisciplinary") |>
  pivot_longer(cols = c(ta_oa, no_ta), names_to = "oa_type", values_to = "oa") |> 
  ggplot(aes(cr_year, oa / jn_all, fill = oa_type)) +
  geom_boxplot(varwidth = TRUE, outlier.shape = NA, notch = TRUE) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent,
                     limits = c(0, .5)) +
  facet_grid(top_level ~ publisher)

subject_short |>
  ungroup() |>
  filter(oa > 0) |>
  select(-oa) |>
  pivot_longer(cols = c(ta_oa, no_ta), names_to = "oa_type", values_to = "oa") |> 
  filter(top_level != "Multidisciplinary") |>
  mutate(publisher = as.character(publisher)) |>
  group_by(publisher, top_level, cr_year) |>

subject_short |>
  ungroup() |>
  filter(oa > 0) |>
  select(-oa) |>
  pivot_longer(cols = c(ta_oa, no_ta), names_to = "oa_type", values_to = "oa") |> 
  filter(top_level != "Multidisciplinary") |>
  mutate(publisher = as.character(publisher)) |>
  group_by(publisher, top_level, cr_year) |>
  summarise(nn = wilcox.test(oa ~ oa-type)$p.value))

subject_short |>
 ungroup() |>
filter(oa > 0) |>
select(-oa) |>
 pivot_longer(cols = c(ta_oa, no_ta), names_to = "oa_type", values_to = "oa") |> 
  mutate(oa_prop = oa / jn_all) |>
 filter(top_level != "Multidisciplinary") |>
 mutate(publisher = as.character(publisher)) |>
  group_by(publisher, top_level, cr_year) |>
  summarise(p_value = wilcox.test(oa ~ oa_type, paired = TRUE)$p.value) |> View()

preprints %>%
  filter(posted_date >= analysis_start,
         posted_date <= analysis_end) %>%
  with(., wilcox.test(n_versions ~ covid_preprint))

```
